### 2021-10-14 paper/paper.tex
cite paper
changes

% If you use BibTeX in apalike style, activate the following line:
%\bibliographystyle{apalike}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{natbib}
\usepackage[inline]{enumitem}
\usepackage{tikz}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{mwe}

@@

\theoremstyle{definition}
\newtheorem{problem}{Problem}
\newtheorem{defn}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{fact}{Fact}

@@

\renewcommand{\star}[1]{{#1}^{*}}
\newcommand{\trans}[1]{{#1}^{T}}
\newcommand{\loss}{\ell}
\newcommand{\aaa}{\mathbf a}
\newcommand{\vv}{\mathbf v}
\newcommand{\uu}{\mathbf u}
\newcommand{\w}{\mathbf w}
\newcommand{\x}{\mathbf x}
\newcommand{\y}{\mathbf y}
\newcommand{\lone}[1]{{\lVert {#1} \rVert}_1}
\newcommand{\ltwo}[1]{{\lVert {#1} \rVert}_2}
\newcommand{\lp}[1]{{\lVert {#1} \rVert}_p}
\newcommand{\linf}[1]{{\lVert {#1} \rVert}_\infty}
\newcommand{\lF}[1]{{\lVert {#1} \rVert}_F}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Tree Loss}

@@

Assume we are solving a standard multi-class classification problem with $k$ classes and $d$ input feature dimensions.
The cross entropy loss (also called the logistic loss) is the standard loss function for this problem.
It is defined to be
\begin{equation}
    \label{eq:xentropy}
    \ell(W;(\x,y)) = - \log \frac {\exp(-\trans\w_y \x)}{\sum_{j=1}^k \exp(-\trans \w_j \x)}
\end{equation}
where for each class $i\in[k]$,
$\w_i : \R^d$ is the parameter vector associated with class $i$;
the variable $W : \R^{k \times d} = (\w_1; \w_2; ...; \w_k)$ is the full parameter matrix;
$\x : \R^d$ is the feature vector;
and $y \in [k]$ is the class label.

@@

The tree loss works by reparameterizing the weight matrix for the cross entropy loss.
Intuitively, our goal is to enforce that semantically similar labels will have similar weight vectors.

The reparameterization is defined according to a \emph{label tree} structure that captures the similarity between labels.
Every class is represented by a leaf node in the label tree,
and internal nodes represent \emph{pseudo-classes} that group the existing classes into a logical hierarchy.
At this point, we assume that the label tree structure is given a priori, and in Section \ref{sec:bias-var} we discuss methods for generating this structure and their tradeoffs.
Figure \ref{fig:labeltree} shows an example label tree for a classification problem with 9 classes.

\begin{figure}
    \resizebox{\columnwidth}{!}{
\begin{tikzpicture}
    [ level distance=1.5cm
    , level 1/.style={sibling distance=3cm}
    , level 2/.style={sibling distance=1.5cm}
    ]
\node {\textit{pseudo1}}
    child {node {\textit{pseudo2}}
      child {node {\texttt{boat}}}
      child {node {\texttt{plane}}}
      child {node {\textit{pseudo3}}
        child {node {\texttt{car}}}
        child {node {\texttt{truck}}}
        child {node {\texttt{van}}}
        child {node {\texttt{motorcycle}}}
      }
      }
    child {node {\texttt{building}}}
    child {node {\textit{pseudo4}}
      child {node {\texttt{cat}}}
      child {node {\texttt{dog}}}
    };
\end{tikzpicture}
}
    \caption{\fixme{caption}}
    \label{fig:labeltree}
\end{figure}

We parameterize the weight vectors from the label tree as follows.
For each leaf node $i$ (equivalently for each class $i$),
we let the sequence $P_i$ denote the path from the leaf node to the root.
For example, using the label tree above,
we have

\begin{equation}
\begin{split}
    P_{\texttt{truck}} &= (\texttt{truck}, \textit{pseudo3}, \textit{pseudo2}, \textit{pseudo1} ) \\
    P_{\texttt{cat}} &= (\texttt{cat}, \textit{pseudo4}, \textit{pseudo1} )
\end{split}
\end{equation}

% \end{equation}

Next, we associate a vector $\vv_i$ to each leaf and internal node $i$ in the tree structure.
In particular, there is a vector $\vv_i$ for each class and for each pseudo-class.
Finally, we reparameterize the weights for the cross entropy loss as the sum of the vectors along the paths.
Specifically, we set the weight vector for class $j$ to be 
\begin{equation}
    \label{eq:covertree-reparam}
    \w_j = \sum_{k\in P_j} \vv_k
    .
\end{equation}
Substituting \eqref{eq:covertree-reparam} into \eqref{eq:xentropy} gives us the final tree loss
\begin{equation}
    \ell(V;(\x,y)) = - \log \frac {\exp(-\sum_{k\in P_y}\trans\vv_k \x)}{\sum_{j=1}^k \exp(- \sum_{k\in P_j}\trans\vv_k \x)}
\end{equation}
where $V=(\vv_1,...,\vv_k)$,
and our goal is to minimize $\ell$ with respect to $V$.
Note that this loss can easily be computed in libraries like pytorch or tensorflow with the standard cross entropy loss functions.
The only change needed is to define the loss in terms of the sum of vectors $V$ instead of a variable tensor $W$.

\subsection{Related loss functions}

% \fixme{Read the papers mentioned above.  Verify that you agree with the things I wrote.  Then, write a detailed mathematical description of the HSM loss and contrast it with the cover tree loss.}

% \fixme{Read 5-10 additional papers about the hierarchical softmax.  For each paper, write a paragraph describing the paper's contributions and how it relates to the cover tree loss.  For the final paper we will reference these papers.  We will have to trim down the notes that you write for this task, but it's useful when writing the paper to have detailed notes about references.}

@@

It is well known that the cross entropy loss has generalization error $O(\sqrt{k/n})$.
Our main theoretical result states that the tree loss significantly improves this generalization error down to $O(\sqrt{1/n})$.
This is a shocking improvement since it implies that the generalization error is independent of the number of classes in our problem.

Our analysis relies on simple properties of stochastic gradient descent,
and we view the simplicity of our analysis as a major strength of the tree loss.
In this section, we first review standard results of stochastic gradient needed for our analysis.
Then we formally state and prove our main result.
We conclude with a discussion about the bias-variance tradeoff.

@@

SGD is a standard algorithm for optimizing convex losses,
and it has many variations.
Our analysis uses the results from \citet{shalev2014understanding} which are based off of a version of SGD that uses parameter averaging.
All variants of SGD have similar results, however, and we refer the reader to the above reference for a detailed discussion of these variants and their algorithmic details.

The main result we use is a bound on the convergence rate of SGD for convex Lipschitz functions.
We reproduce Theorem 14.8 of \citet{shalev2014understanding} in the theorem below.

\begin{theorem}
    \label{thm:sgd}
    Let $B,\rho>0$.
    Let $f$ be a convex function and let $\star W = \argmin_{W : \lF{W}\le B} f(W)$.
    Assume that SGD is run for $T$ iterations with $\eta = \sqrt{\frac{B^2}{\rho^2 T}}$.
    Assume that for all $t$, $\ltwo{\nabla f(W_t)} \le \rho$.
    Then,
    \begin{equation}
        \E f(\bar W) - f(\star W) \le \frac{B\rho}{\sqrt{T}}
        .
    \end{equation}
\end{theorem}

The following theorem bounds the generalization error of the cross entropy loss for multiclass classification problems.
It is an immediate consequence of Theorem \ref{thm:sgd} above when $f$ is the cross entropy loss.

\begin{theorem}
\label{theorem:xentropy}
Assume that for all $i\in[n]$, $\ltwo{\x_i} \le \rho$,
    and that for each class $i\in[k]$, $\ltwo{\w_i}\le B$.
Then the parameter vector $\bar W$ estimated from one epoch of SGD satisfies
\begin{equation}
    \E L_D(\bar W) - L_D(W^*) \le \frac {\sqrt kB\rho}{\sqrt n}
\end{equation}
where $W^* = \argmin L_D(W)$.
\end{theorem}

Intuitively, this theorem states that the generalization error of the cross entropy loss shrinks at the rate $O(\sqrt{k/n})$.

%\begin{theorem}
    %Let $B,\rho>0$.
    %Let $f$ be a convex function and let $\star \w = \argmin_{\w : \ltwo{\w}\le B} f(\w)$.
    %Assume that SGD is run for $T$ iterations with $\eta = \sqrt{\frac{B^2}{\rho^2 T}}$.
    %Assume that for all $t$, $\ltwo{\nabla f(\w_t)} \le \rho$ with probability 1.
    %Then,
    %\begin{equation}
        %\E f(\bar \w) - f(\star \w) \le \frac{B\rho}{\sqrt{T}}
        %.
    %\end{equation}
%\end{theorem}
%
%When analyzing stochastic gradient descent using the cross entropy loss,
%it is common to assume that the parameter space is bounded.
%Under the standard parameterization, we have that the weight vector for each class satisfies $\ltwo{\w_j} \le B$.
%It immediately follows that $\lF{W} \le \sqrt{k}B$,
%and from this and convergence results of SGD it follows that the generalization error is $O(\sqrt{k}B/n)$.
%(See Theorem \ref{theorem:xentropy} above.)
%
%\begin{defn}
    %A learning problem is called convex-lipschitz-bounded with parameters $\rho,B$ if
    %\begin{enumerate}
        %\item $\ltwo{\w} \le B$
        %\item the loss
    %\end{enumerate}
%\end{defn}
%
%\begin{corollary}[Corollary 14.12 \citep{shalev2014understanding}]
    %Consider a convex-Lipschitz-bounded learning problem with parameters $\rho, B$.
    %Then for every $\epsilon>0$, if we run the SGD method for minimizing $L_\mathcal D(W)$ with a number of iterations
        %$T \ge \frac{B^2\rho^2}{\epsilon^2}$
    %and with
        %$\eta = \sqrt{\frac{B^2}{\rho^2 T}}$
    %then the output of SGD satisfies
    %\begin{equation}
        %\E L_\mathcal D (\bar W) - L_\mathcal D(\star W) \le \epsilon
    %\end{equation}
%\end{corollary}
%
%\subsection{Background}

%In this section, we first introduce more notation and definitions needed for our analysis,
%then we recall a standard result of stochastic gradient descent.
%We follow the notation of \citet{shalev2014understanding}'s textbook as closely as possible.
%
%Let $\mathcal D$ be a distribution over the input data points and $S \sim D^m$ be a set of $m$ data points sampled from this distribution.
%Then we define the true loss as
%\begin{equation}
    %L_\mathcal D(W) = \E_{(\x,y)\sim\mathcal D} \ell(W; \x, y)
%\end{equation}
%and the sample loss as
%\begin{equation}
    %L_S(W) = \frac 1 m \sum_{i=1}^m \ell(W; \x_i, y_i)
    %.
%\end{equation}
%We refer to the difference between these two losses as the generalization error.
%\begin{equation}
    %\left\lvert L_\mathcal D(W) - L_S(W) \right\rvert.
%\end{equation}
%
%Given a distribution of data points $\mathcal D$, the optimal parameter matrix is defined as
%\begin{equation}
    %\star W = \argmin_{W} \E_{(\x, y)} \ell(W; (\x, y))
%\end{equation}
%Given a dataset of $n$ points, we can estimate the parameter 
%\begin{equation}
    %\hat W = \argmin_{W} \frac1n\sum_{i=1}^n \ell(W; (\x_i, y_i))
%\end{equation}
%The main theoretical result for the cross entropy loss is that for a dataset with $n$ points,
%the generalization error is $O(\sqrt{k/n})$.
%There are many theorems that establish this result in different contexts.
%The following is an example that relies on properties of SGD.
%\begin{theorem}
%\label{theorem:xentropy}
%Assume that for all $i\in[n]$, $\ltwo{x_i} \le \rho$,
    %and that for each class $i\in[k]$, $\ltwo{\w_i}\le B$.
%Then the parameter vector $\bar W$ estimated from one epoch of SGD satisfies
%\begin{equation}
    %\E L_D(\bar W) - L_D(W^*) \le \frac {\sqrt cB\rho}{\sqrt n}
%\end{equation}
%where $W^* = \argmin L_D(W)$.
%\end{theorem}
% \begin{proof}
%     \fixme{Write a proof of this result using Theorem 14.12 of Shalev-Shwartz and Ben-David and the properties of the cross entropy loss from the last set of notes.}
%     \begin{align}
%     {\| W \|^2_F}&=\sum_{i=1}^k{\| W_i\|^2}\\
%                 &\le\sum_{i=1}^k{B^2}\\
%                 &\le cB^2\\
%   \| W \|_F  &\le \sqrt k B 
%     \end{align}
% Then we get:
%     \begin{equation}
%     l(\bar W)-l(W^*)\le\frac {\sqrt cB\rho}{\sqrt n}
%     \end{equation}
%   Take expectation of both side of equation(3):
%   \begin{equation}
%     \E L_D(\bar W) - L_D(W^*) \le \frac {\sqrt cB\rho}{\sqrt n}
% \end{equation}
% \end{proof}

%There are many other loss functions that have been proposed for multilabel classification.
%In this section, we review these alternatives before presenting our proposed the tree loss.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Bounds using L2 regularization}
%
%Recall that when we apply L2 regularization,
%we are minimizing the objective function
%\begin{equation}
    %f(W) = \tfrac 1 m \sum_{i=1}^m \loss(W; (\x_i,y_i)) + \tfrac\lambda2\lF{W}^2
    %.
%\end{equation}
%This function is $\lambda$-strongly convex in $W$.
%(Why?)
%Then Theorem 14.11 tells us that after performing $T$ iterations of SGD, we have that
%\begin{equation}
    %\label{eq:f-bound}
    %\E f(\bar W) - f(W^*) \le \frac{4\rho^2}{\lambda T} ( 1 + \log T)
%\end{equation}
%where $W^* = \argmin f(W)$.
%If we substitute $f(W)=L_D(W)+\tfrac\lambda 2\lF{W}^2$ into Equation \ref{eq:f-bound} above,
%then we get
%\begin{align}
    %\E L_D(\bar W) - L_D(W^*) 
    %&\le \frac{4\rho^2}{\lambda T} ( 1 + \log T) + \tfrac\lambda2\lF{W^*}^2 - \tfrac\lambda2\lF{\bar W}^2
    %\\
    %&\le \frac{4\rho^2}{\lambda T} ( 1 + \log T) + \frac{\lambda k B^2}{2}
    %.
%\end{align}

>>>>

% If you use BibTeX in apalike style, activate the following line:
%\bibliographystyle{apalike}
%\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{bm}
\usepackage[round]{natbib}
\usepackage[inline]{enumitem}
\usepackage{tikz}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{mwe}

\usepackage{array}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\newcolumntype{C}[1]{>{\PreserveBackslash\centering}p{#1}}
\newcolumntype{R}[1]{>{\PreserveBackslash\raggedleft}p{#1}}
\newcolumntype{L}[1]{>{\PreserveBackslash\raggedright}p{#1}}

@@

%\theoremstyle{definition}
\newtheorem{assumption}{Assumption}
\newtheorem{problem}{Problem}
\newtheorem{defn}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{fact}{Fact}

@@

\newcommand{\parent}[1]{\texttt{parent}({#1})}

\renewcommand{\star}[1]{{#1}^{*}}
\newcommand{\bad}[1]{{#1}^{\textit{bad}}}
\newcommand{\trans}[1]{{#1}^{T}}
\newcommand{\loss}{\ell}
\newcommand{\aaa}{\mathbf a}
\newcommand{\vv}{\mathbf v}
\newcommand{\uu}{\mathbf u}
\newcommand{\w}{\mathbf w}
\newcommand{\x}{\mathbf x}
\newcommand{\y}{\mathbf y}
\newcommand{\lone}[1]{{\lVert {#1} \rVert}_1}
\newcommand{\ltwo}[1]{{\lVert {#1} \rVert}_2}
\newcommand{\lp}[1]{{\lVert {#1} \rVert}_p}
\newcommand{\linf}[1]{{\lVert {#1} \rVert}_\infty}
\newcommand{\lF}[1]{{\lVert {#1} \rVert}_F}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\resizebox{\columnwidth}{!}{
\begin{tikzpicture}
    [ level distance=1.5cm
    , level 1/.style={sibling distance=3cm}
    , level 2/.style={sibling distance=1.5cm}
    ]
\node[draw, rounded corners=0.1in, inner sep=0.1in] at (-1.5in,0){\textbf{$U$-Tree}};
\node {\texttt{truck}}
    child {node {\texttt{boxer}}
      child {node {\texttt{tiger}}}
      child {node {\texttt{skunk}}}
      child {node {\texttt{boxer}}
        child {edge from parent[draw=none]} % Added
        child {node {\texttt{bulldog}}}
        child {node {\texttt{boxer}}}
        child {node {\texttt{husky}}}
        child {node {\texttt{sheepdog}}}
      }
      child {node {\texttt{bear}}}
      child {edge from parent[draw=none]} % Added
      }
    child {node {\texttt{truck}}
      child {edge from parent[draw=none]} % Added
      child {node {\texttt{truck}}}
      child {node {\texttt{bus}}}
    }
    child {node {\texttt{toaster}}}
    ;

\end{tikzpicture}
}
~~~
\resizebox{\columnwidth}{!}{
\begin{tikzpicture}
    [ level distance=1.5cm
    , level 1/.style={sibling distance=3cm}
    , level 2/.style={sibling distance=1.5cm}
    ]
\node[draw, rounded corners=0.1in, inner sep=0.1in] at (-1.5in,0){\textbf{$V$-Tree}};
\node {\textit{pseudo1}}
    child {node {\textit{pseudo2}}
      child {node {\texttt{tiger}}}
      child {node {\texttt{skunk}}}
      child {node {\textit{pseudo3}}
        child {edge from parent[draw=none]} % Added
        child {node {\texttt{bulldog}}}
        child {node {\texttt{boxer}}}
        child {node {\texttt{husky}}}
        child {node {\texttt{sheepdog}}}
      }
      child {node {\texttt{bear}}}
      child {edge from parent[draw=none]} % Added
      }
    child {node {\textit{pseudo4}}
      child {edge from parent[draw=none]} % Added
      child {node {\texttt{truck}}}
      child {node {\texttt{bus}}}
    }
    child {node {\texttt{toaster}}}
    ;

\end{tikzpicture}
}
    \caption{
        Example label tree structures for a subset of 10 ImageNet classes.
        The $U$-tree has class labels that repeat at multiple levels,
        and the $V$-tree introduces ``pseudoclasses''.
        The pseudoclass \textit{pseudo3} represents the class of ``dogs'',
        and the pseudoclass \textit{pseudo2} represents the class of ``animals''.
    }
    \label{fig:labeltree}
\end{figure*}

\section{The Tree Loss}

This section first formally defines our problem setting,
then formally defines the tree loss.
We conclude with a comparison between the tree loss and related loss functions.

@@

We consider the multi-class classification setting with $k$ classes and $d$ input feature dimensions.
The cross entropy loss is the standard loss function for this setting.
It is defined to be
\begin{equation}
    \label{eq:xentropy}
    \ell(W;(\x,y)) = - \log \frac {\exp(-\trans\w_y \x)}{\sum_{j=1}^k \exp(-\trans \w_j \x)}
\end{equation}
where for each class $i\in[k]$,
$\w_i : \R^d$ is the parameter vector associated with class $i$;
the variable $W : \R^{k \times d} = (\w_1; \w_2; ...; \w_k)$ is the full parameter matrix;
$\x : \R^d$ is the input feature vector;
and $y \in [k]$ is the input class label.

@@

The tree loss reparameterizes the weight matrix used in the cross entropy loss in order to enforce that semantically similar labels will have similar weight vectors.
The reparameterization is defined according to a \emph{label tree} structure that captures the similarity between labels.
And we propose two variants of the tree loss called the $U$ and $V$ variants that use slightly different label tree structures.
The main purpose of the $U$-tree structure is to simplify the analysis and exposition of the $V$-tree structure,
and we show later in our theoretical results that the $V$-tree is strictly better than the $U$-tree.
Figure \ref{fig:labeltree} shows examples of both.

\subsubsection{The $U$-Tree Loss}

In a $U$-tree, every node is a class and the class of all internal nodes matches the class of one of its children.
For each class $i$, we define $\parent{i}$ to be the parent of the upper-most node for class $i$,
and the $\texttt{parent}$ of the root is the 0 vector..
For example, in the example of Figure \ref{fig:labeltree},
we have
\begin{equation*}
\begin{split}
    \parent{\texttt{husky}} &= \texttt{boxer} \text{, and}\\
    \parent{\texttt{bear}} &= \texttt{boxer} \text{, and}\\
    \parent{\texttt{boxer}} &= \texttt{truck} .\\
\end{split}
\end{equation*}
We associate for each class $i$ a new parameter vector $\uu_i$ and set
\begin{equation}
    \w_i = \w_{\parent{i}} - \uu_i,
    \label{eq:uu}
\end{equation}
and we define the parameter matrix $U : \mathbb R^{k\times d}$ to be $(\uu_1;...;\uu_k)$.
The $U$-tree loss is then defined by substituting \eqref{eq:uu} into \eqref{eq:xentropy} to get
\begin{equation*}
    \ell(U;(\x,y)) = - \log \frac {\exp(-(\trans\uu_{\parent{y}}-\trans\uu_y) \x)}{\sum_{j=1}^k \exp(-(\trans\uu_{\parent{j}}-\trans \uu_j) \x)}
\end{equation*}
We use the same $\ell$ notation for the standard cross entropy loss function and the $U$-Tree loss because the function is exactly the same;
the only difference is how we represent the parameters.

\subsubsection{The $V$-Tree Loss}

The $V$-tree is constructed from the $U$-tree by replacing all non-leaf nodes with ``pseudoclasses''.
Let $k'$ be the total number of nodes in the $V$-tree;
so $k'$ equals $k$ plus the number of internal nodes.
For each class $i$, we let the sequence $P_i$ denote the path from the leaf node to the root.
For example, using the $V$-tree in Figure \ref{fig:labeltree},
we have
\begin{equation*}
\begin{split}
    P_{\texttt{sheepdog}} &= (\texttt{sheepdog}, \textit{pseudo3}, \textit{pseudo2}, \textit{pseudo1} ) \\
    P_{\texttt{truck}} &= (\texttt{truck}, \textit{pseudo4}, \textit{pseudo1} )
\end{split}
\end{equation*}
Next, we associate a vector $\vv_i$ to each node $i$ in the tree structure.
In particular, there is a vector $\vv_i$ for each class and for each pseudo-class,
so there are more parameter vectors for the $V$-tree loss than either the $U$-tree loss or the standard cross entropy loss.

Finally, we reparameterize the weights for the cross entropy loss as the sum of the vectors along the paths.
Specifically, we set the weight vector for class $i$ to be 
\begin{equation}
    \label{eq:covertree-reparam}
    \w_j = \sum_{j\in P_i} \vv_i
    .
\end{equation}
The $V$-tree loss is defined by substituting \eqref{eq:covertree-reparam} into \eqref{eq:xentropy} to get
\begin{equation}
    \ell(V;(\x,y)) = - \log \frac {\exp(-\sum_{k\in P_y}\trans\vv_k \x)}{\sum_{j=1}^k \exp(- \sum_{k\in P_j}\trans\vv_k \x)}
    .
\end{equation}
We continue to use the $\ell$ function to represent both the standard cross entropy loss and the $V$-tree loss function to emphasize that these are the same loss functions,
just with different parameterizations of the weight matrix.

\subsection{Intuition}

The intuition behind the $V$-tree reparameterization is that whenever we perform a gradient update on a data point with class $i$,
we will also be ``automatically'' updating the parameter vectors of similar classes.
To see this, note that when we have a \texttt{sheepdog} data point,
we will perform gradient updates on all $\vv_i$ in $P_\texttt{sheepdog}$; i.e. $\vv_{sheepdog}$, $\vv_{\textit{pseudo3}}$, $\vv_{\textit{pseudo2}}$, and $\vv_{\textit{pseudo1}}$.
This will cause the $\w_{\texttt{husky}}$ and $\w_{\texttt{bear}}$ parameters (among many others) to update because they also depend on the pseudoclass vectors $\vv_{\textit{pseudo3}}$ and $\vv{_\textit{pseudo2}}$.
Because $P_\texttt{husky}$ has a larger overlap with $P_\texttt{sheepdog}$ than $P_\texttt{bear}$,
the parameters of this more similar class will be updated more heavily.

\subsection{Implementation Notes}

Both the $U$-tree and $V$-tree losses are easy to implement in practice using the built-in cross entropy loss function in libraries like PyTorch \citep{NEURIPS2019_9015} or Tensorflow \citep{tensorflow2015-whitepaper}.
The only change needed is to define the loss in terms of the sum of vectors $V$ instead of a variable tensor $W$.
This ensures that the loss is implemented efficiently and is numerically stable.

In practice, the $V$-tree loss is slightly slower than the standard cross entropy loss and $U$-tree losses due to the additional summation over the paths.
In our experiments, we found this this to cause the $V$-tree loss to be about twice as slow.

@@

It is well known that when using the standard cross entropy loss,
SGD converges at a rate of $O(\sqrt{k/n})$.
Our theoretical results asymptotically improve this convergence rate.

With an appropriately constructed label tree,
we show convergence at a rate of $O(\sqrt{\log k/n})$ or $O(\sqrt{k^{1-1/c}/n})$,
where $c$ is a measure of difficulty of the problem.
But what happens if the tree is poorly constructed?
Remarkably, nothing bad.
We also show that for arbitrarily bad label tree structures,
the $U$ and $V$ losses maintain a $O(\sqrt{k/n})$ convergence rate.
This suggests that there is essentially no downside to using the tree loss parameterizations.

\fixme{
Our analysis relies on simple properties of stochastic gradient descent,
and we view the simplicity of our analysis as a major strength of the tree loss.
In this section, we first review standard results of stochastic gradient descent needed for our analysis.
Then we formally state and prove our main result.
}

@@

SGD is a standard algorithm for optimizing convex losses,
and it has many variations.
Our analysis uses the results from Chapter 14 of \citet{shalev2014understanding} which are based off of a version of SGD that uses parameter averaging.\footnote{
Other variants of SGD have similar results,
and our results are easy to generalize to other SGD variants.
We refer the reader to \citet{shalev2014understanding} for a discussion of these variants.}
In this section we recall these results,
occasionally changing the names of variables to avoid conflicts with variable names used elsewhere in this paper.

We assume we are optimizing a function $f : \Theta \to \mathbb R$,
and that we are only given noisy samples of the function and its gradient.
We let $\theta^{t}$ denote the parameter value at iteration $t$,
and $\eta : \mathbb R$ be the step size.
Then the parameter update for step $t$ is defined to be
\begin{equation}
    \theta^{(t+1)} = \theta^{(t)} - \eta \nabla f(\theta^{(t)}).
\end{equation}
When using parameter averaging, the final model parameters are defined to be
\begin{equation}
    \bar\theta = \frac1T \sum_{i=1}^T \theta^{(t)}
    .
\end{equation}
Parameter averaging is not often used in practice because of the memory requirements,
but we use it here due because theorems using parameter averaging are typically much easier to state and prove.
Equivalent results without parameter averaging typically require and additional $\log T$ factor in their convergence rates\citep{}.

The main result we rely on is a bound on the convergence rate of SGD for convex Lipschitz functions.
We reproduce Theorem 14.8 of \citet{shalev2014understanding} in the theorem below.

\begin{theorem}
    \label{thm:sgd}
    Let $\beta,\rho>0$.
    Let $f$ be a convex function and let $\star \theta = \argmin_{\theta : \ltwo{\theta}\le \beta} f(\theta)$.
    Assume that SGD is run for $T$ iterations with $\eta = \sqrt{\frac{\beta^2}{\rho^2 T}}$.
    Assume that for all $t$, $\ltwo{\nabla f(\theta^{(t)})} \le \rho$.
    Then,
    \begin{equation}
        \E f(\bar \theta) - f(\star \theta) \le \frac{\beta\rho}{\sqrt{T}}
        .
    \end{equation}
\end{theorem}

\subsection{SGD and Statistical Learning}

We now introduce more statistical learning theory notation,
then apply Theorem \ref{thm:sgd} above to bound the rate of learning.

Define the true loss of our learning problem as
\begin{equation}
    L_D(A) = \E_{(\x,y)\sim D} \ell(A; (\x, y))
\end{equation}
where $A \in \{W, U, V\}$ is a parameterization of the cross entropy loss.
The optimal parameter vector is defined to be the minimum of the true loss:
\begin{equation}
    \star A = \argmin_{A} L_D(A).
\end{equation}
In order to apply Theorem \ref{thm:sgd} to a learning problem,
we make the following assumption about our the size of our data points:
\begin{assumption}
    \label{ass:lip}
    For each feature vector $\x$ in the data set, 
        $\ltwo{\x} \le \rho$
        .
\end{assumption}
Under this assumption, the cross entropy loss $\ell$ is $\rho$-Lipschitz for all parameterizations,
and the true loss $L_D$ is therefore also $\rho$-Lipschitz.

Applying Theorem \ref{thm:sgd} with $f=L_D$ gives
\begin{corollary}
    \label{ref:cor:A}
    Under Assumption \ref{ass:lip}, we have that for any parameterization $A$ of the cross entropy loss,
    \begin{equation}
        \E L_D(\bar A) - L_D(\star A) \le \frac{\lF{\star A}\rho}{\sqrt n}.
        \label{eq:Aconv}
    \end{equation}
\end{corollary}

We can now bound the convergence of the standard cross entropy loss and the $U$ and $V$-tree losses by bounding the Frobenius norms of their parameter matrices.

In order to recover the standard convergence rate of multiclass classification, we make the following assumption.
\begin{assumption}
    \label{ass:B}
    For each class $i$, the optimal parameter vector $\star\w_i$ satisfies
        $\ltwo{\star\w_i} \le B$
        .
\end{assumption}
Note that this assumption is not limiting because the value of $B$ need not be known during the optimization procedure.

It follows that
\begin{equation}
    \lF{\star W}^2 = \sum_{i=1}^k \ltwo{\star\w_i} \le kB^2.
    \label{eq:starW}
\end{equation}
Substituting Eq \eqref{eq:starW} into \eqref{eq:Aconv} gives the following bound.

\begin{theorem}
\label{theorem:xentropy}
    Under assumptions \ref{ass:lip} and \ref{ass:B} above,
then the parameter vector $\bar W$ estimated from one epoch of SGD satisfies
\begin{equation}
    \E L_D(\bar W) - L_D(W^*)
    \le \frac {\sqrt kB\rho}{\sqrt n}
    .
\end{equation}
\end{theorem}
\fixme{
It is important to note that this result above is an upper bound and not a lower bound.
This $O(\sqrt{k/n})$ convergence rate matches the convergence rate implied by the Nataranj dimension \citep{}.
We are not aware of any lower bounds for SGD in the Lipschitz setting.
\citet{nguyen2018tight,jentzen2020lower} study the strongly convex case and find that the upper bounds match the lower bounds.
}

\subsection {Bad Tree Structures}

\begin{lemma}
    \ref{lemma:starV}
    We have that for all $V$-trees derived from a $U$-tree,
    \begin{equation}
        \lF{\star V} \le \lF{\star U}.
    \end{equation}
\end{lemma}
\begin{proof}
    \fixme{}
\end{proof}

The following lemma shows that even for arbitrarily bad tree structures,
the bound on the parameter spaces cannot be too bad.

\begin{lemma}
    \ref{lemma:starU}
    Under assumption \ref{ass:B},
    the following bound holds for all $U$-tree structures:
    \begin{equation}
        \lF{\star U} \le 2\sqrt{k}B.
    \end{equation}
\end{lemma}
\begin{proof}
    We have
    \begin{align}
        \lF{\star U}^2 
        &= \sum_{i=1}^k \ltwo{\star\uu_i}^2 \\
        &= \sum_{i=1}^k \ltwo{\star\w_i - \star\w_{\parent{i}}}^2 \\
        &\le \sum_{i=1}^k \left(\ltwo{\star\w_i} + \ltwo{\star\w_{\parent{i}}} \right)^2 \\
        &\le \sum_{i=1}^k (2B)^2 \\
        &\le k (2B)^2
        .
    \end{align}
\end{proof}

\subsection{Good Tree Structures}

\begin{assumption}
We are given a distance metric $d$ over the labels such that for all labels $i$ and $j$,
\begin{equation}
    \frac 1 \lambda d(i,j)
    \le \ltwo{\star \w_i - \star \w_j}
    \le \lambda d(i, j).
\end{equation}
We denote by $c$ the doubling dimension of the metric $d$.
\end{assumption}

\begin{lemma}
    \ref{lemma:main}
    When $c\le1$, we have that
    \begin{equation}
        \lF{\star U} \le \tfrac{1}{\sqrt2}\lambda B \sqrt{\log_2 k}.
        \label{eq:c<=1}
    \end{equation}
    When $c>1$, we have that
    \begin{equation}
        \lF{\star U} \le \sqrt{5}\lambda B \sqrt{k^{(1-1/c)}}.
        \label{eq:c>1}
    \end{equation}
\end{lemma}
\begin{proof}
%Construct a cover tree over the distance metric $d$.
%For each class $i$, define the vector
%\begin{equation}
    %\uu_i =
    %\begin{cases}
        %\w_i & \text{if $i$ is the root of the cover tree} \\
        %\w_i - \w_{\parent{i}} & \text{otherwise} \\
    %\end{cases}
    %.
%\end{equation}
%Define the vector $\star\uu_i$ similarly by replacing all the $\w_i$ with $\star\w_i$ above.
    First, we review some simple properties of the cover tree structure that will be needed in our proof.
    Let $q$ denote the maximum number of children of any node in the cover tree.
    A standard property of cover trees is that $q \le (4^c)$ \citep[Lemma 25]{izbickithesis}.
    Let $Q_i$ denote the number of nodes at level $i$ in the cover tree.
    Then we have that $|Q_i| \le q^i.$
    Finally, let $r$ be the height of the cover tree.
    Without loss of generality, we analyze the bound when the cover tree is perfectly balanced and so $r = \log_q k = \log_2 k / (2c)$.
